The vanishing redshift as a probe of reionization

### 2020-02-20 paper/ms.tex
Fix previous commit
Update author order

\newcommand{\zst}{\ensuremath{z_{\star}}}
\newcommand{\E}[1]{\mathrm{E}[#1]}
\newcommand{\Var}[1]{\mathrm{Var}[#1]}
\newcommand{\Cov}[2]{\mathrm{Cov}[#1,#2]}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[The transition redshift as an EoR probe]{The transition redshift as a probe of reionization}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[A. Beane et al.]{
Angus Beane,$^{1}$\thanks{E-mail: angus.beane@cfa.harvard.edu}
Adam Lidz,$^{2}$
Kana Moriwaki,$^{3}$
and others
\\
% List of institutions
$^{1}$Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
$^{2}$Department of Physics \& Astronomy, University of Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104, USA\\
$^{3}$Department of Physics, The University of Tokyo, 7-3-1 Hongo, Bunkyo, Tokyo 113-0033, Japan
% $^{2}$Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, NY 10010, USA\\
}

@@

% Abstract of the paper
\begin{abstract}
The early stages of the Epoch of Reionization (EoR), probed by the 21~cm line,
are sensitive to the detailed formation of the first galaxies. An important
transition occurs on large scales at $z\sim10$, when the 21~cm field changes
from being positively correlated with the density field to being negatively
correlated. The redshift at which this occurs (the ``transition redshift'') is
in principle measurable if the 21~cm field can be cross-correlated with a
tracer of the density field (such as galaxies or a line intensity map) at that
high redshift. Since the transition redshift is derived from a
cross-correlation, it naively depends on the structure of the 21~cm field and
the tracer field. However, we show that on large scales the transition
redshift is independent of the structure of the tracer field. We verify this
in simulations by assuming a simple model for a mock Ly$\alpha$ intensity
mapping survey. By considering a joint measurement between the Square
Kilometre Array and the Cosmic Dawn Intensity Mapper, we show that the
transition redshift is measurable up until about $z\sim x$. Being derived from
a cross-correlation, the transition redshift is less impacted by foreground
contamination currently inhibiting 21~cm auto-spectrum measurements.
Therefore, the transition redshift is an interesting probe of the early stages
of the EoR, and its detectability and interpretability deserve further study.
\end{abstract}

@@

@@

This cross-correlation was first studied with the CO rotational lines
\citep{2011ApJ...741...70L}, and then the [\CII] \citep{2012ApJ...745...49G}
and Ly$\alpha$ lines \citep{2013ApJ...763..132S}. The possibility of using the
cross-bispectrum the 21~cm and [\CII] fields has been explored
\citep{2018ApJ...867...26B}, along with combining the three cross-spectra with
a 21~cm, [\CII], and [\OIII] overlap \citep{2019ApJ...874..133B}. Because the
transition redshift is typically $\z_{\star} \sim 10$, it is unclear how much
metal enrichment will have occurred by this early time.

In the case of either galaxies or emission lines, it is expected that the two
are positively correlated at the very beginning of the EoR (though the
emission lines may be very weak due to the lack of significant metal
enhancement). This is because the densest regions host the most mass and
therefore 21~cm emission and galaxy populations. However, the densest regions
will also ionize first {\bf cite something}, and after some time 21~cm and
galaxies will become negatively correlated. We refer to the redshift at which
this transition occurs at a fixed $k$ as the ``transition redshift''
$z_\star(k)$, assuming that only one such zero exists. In other words, the
cross-correlation at $k$ vanishes at $z=z_{\star}(k)$.

In this work we investigate to what extent the transition redshift is
dependent upon the galaxy sample used to define it. We assume a galaxy catalog
selected using an [\OIII] emitter survey. In particular, we are interested in
how $z_\star$ changes as we vary the parameters of the [\OIII] luminosity-mass
relation for a fixed halo catalog.

\section{Methods} \label{sec:methods}
\subsection{The transition redshift} \label{ssec:ztran_def}
The transition redshift $\zst(k)$ is defined by a field $x$ which traces the
density field on large scales and the condition that
\beq \label{eq:ztran_def}
P_{21,x}(k, \zst(k)) \equiv 0\text{.}
\eeq
If more than one such \zst{} exists, then we choose the lowest one where
$P_{21,x}$ last transitioned from positive to negative.

@@

\section{Results} \label{sec:results}
\subsection{The transition redshift} \label{ssec:ztran_results}

\begin{figure*}
\includegraphics[width=\textwidth]{plots/aveTb_nf.pdf}
\caption{A summary of the ionization and heating history of our three EoR
models: \hot{}, \fid{}, and \cold{}. \textit{Upper}: the ionization history of
each model, showing that reionization begins ($\avg{x_{\HII}}\sim0.01$) at
$z\sim16$ and ends at $z\sim6$ in each model. In detail the ionization
histories are slightly different, but not drastically. \textit{Lower}: the
heating history of each model, showing the strong differences in the monopole
signal. The transition redshift \zst{} (defined where $P_{21,\delta}(\zst{}) =
0$) for each model is shown as a vertically dashed line. We see that models in
which heating is more efficient have an earlier \zst{}.}
\label{fig:aveTb_nf}
\end{figure*}

@@

We first investigate the dependence of the transition redshift as defined by
the halo field. Fig.~\ref{fig:ztran_vs_k} shows the transition redshift
defined by the 21~cm\,--\,density cross-power spectrum (see
Eq.~\ref{eq:ztran_def}) as a function of $k$. The three models considered in
this work are shown in orange, gray, and teal (\hot{}, \fid{}, and \cold{},
respectively).

Fig.~\ref{fig:ztran_vs_k} shows that \zst{} is only weakly a function of $k$.

\begin{figure}
\includegraphics[width=\columnwidth]{plots/ztran_vs_k.pdf}
\caption{The transition redshift ($\zst$) as defined by cross-correlating with the
density field (solid lines) and all resolved halos (dashed lines, $M\gtrsim
5\times10^9\,\Msun$), for our three different EoR models (\hot{}, \fid{}, and
\cold{}). We see that \zst{} is only a mild function of $k$ for the three
models, with slightly stronger dependence as the heating of the model
decreases. The most dependence comes in the \cold{} model, where \zst{}
changes by $\sim7\,\%$ from $k=0.1$ to $0.6\,\hoverMpc$. {\bf Some words for
why cold and fid cut off at some k}}
\label{fig:ztran_vs_k}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{plots/ztran_vs_Mcut.pdf}
\caption{The transition redshift (\zst) as defined by cross-correlating with
the halo field for our three different EoR models (\hot{}, \fid{}, and
\cold{}) as a function of the cut-off mass. The cross-correlation is only made
between the 21~cm field and all halos with masses $M>M_{\text{cut}}$. One can
see that for the \fid{} and \cold{} models, \zst{} is only mildly a function
of $M_{\text{cut}}$ below a certain cut-off mass ($\sim10^{10.5}\,\Msun$ in
both models). In other words, as long as a galaxy survey is deep enough so as
to resolve a sufficient number of galaxies, the \zst{} is not dependent on the
depth of the survey. The \hot{} model shows more dependence, but this may be
due to sampling noise due to the smaller number of such halos at the higher
\zst{} typical for the \hot{} model.}
\end{figure}

\subsection{Line-intensity mapping} \label{ssec:lim}
We now turn to understanding how \zst{} depends on the parameters of our
generic line-intensity mapping field: the slope ($\alpha$) and lognormal
scatter ($\sigma$). See \S~\ref{ssec:oiii_int_map} for more details of the
construction of the field.

In Fig.~\ref{fig:ztran_vs_alpha_sigma} we plot \zst{} as a function of each of
these parameters\,--\,$\alpha$ and $\sigma$. We see that \zst{} is remarkably
insensitive to the value of either of these parameters, with some small
deviation as a function of $\alpha$. The \hot{}, \fid{}, and \cold{} models
only vary by a total of $1.2\%$, $0.56\%$, $0.52\%$ with respect to $\alpha$,
and $0.070\%$, $0.10\%$, and $0.11\%$ with respect to $\sigma$, respectively.

\begin{figure*}
\includegraphics[width=\textwidth]{plots/ztran_vs_alpha_sigma.pdf}
\caption{The transition redshift (\zst) defined by cross-correlating with a
generic two-parameter intensity mapping field. The parameter $\alpha$ sets the
slope of the luminosity-mass relation and the parameter $\sigma$ sets the
lognormal scatter in this relation (see \S~\ref{ssec:oiii_int_map}). The
\hot{}, \fid{}, and \cold{} models only vary by a total of $1.2\%$, $0.56\%$,
$0.52\%$ with respect to $\alpha$, and $0.070\%$, $0.10\%$, and $0.11\%$ with
respect to $\sigma$, respectively. Clearly \zst{} varies only with the
structure of the 21~cm field, and is thus an excellent probe of reionization.}
\label{fig:ztran_vs_alpha_sigma}
\end{figure*}

\section{Detectability}
Up to this point, we have not been specific about which galactic emission line
one might consider using to measure \zst{}\,--\,in fact, our entire formalism
could be executed using a galaxy survey, if one with sufficient redshift
accuracy and depth could be executed. While it will not be possible for a
traditional galaxy survey to extend to $z\sim10$ in the near future, it may be
possible to execute a survey of line-emitting galaxies, e.g., with
[\OIII]-emitters \citep{2018MNRAS.481L..84M, 2019arXiv190610863M}.

@@

The number of modes is given by,
\beq \label{eq:nmodes}
N_m = \frac{4\pi k^2 \delta k}{V_{\text{fund}}}\text{,}
\eeq
where $\delta k$ is the width of the $k$-bin and $V_{\text{fund}}$ is the
volume of a fundamental mode. We assume a square survey area, and so this is
given by,
\beq \label{eq:fund}
V_{\text{fund}} = \frac{(2 \pi)^3}{L_{\perp}^2 L_{\parallel}}\text{,}
\eeq
where $L_{\perp}$ is the  and $L_{\parallel}$ is the comoving length of the
redshift extent of the bin. The optimal redshift binning is not entirely
clear, since the observational target is not the power spectrum. We assume a
redshift bin of $\Delta z = 0.03$, which is smaller than the expected error we
compute on \zst{}, and so is not an unreasonable choice. We assume a $k$-bin
of fairly large width, from $k=0.1$ to $0.4\,\Mpc^{-1}$.

By expanding $P_{21,\Ha}$ as a linear function about \zst{}, we can estimate the
error on \zst{} by
\beq \label{eq:error_on_zst}
\Var{\zst} = \left(\left.\frac{\partial P_{21,\Ha}}{\partial z}\right|_{\zst}\right)^{-2} \Var{P_{21,\Ha}}\text{,}
\eeq
where ${\partial P_{21,\Ha}}/{\partial z}$ is computed numerically.

>>>>

\newcommand{\zv}{\ensuremath{z_{v}}}
\newcommand{\E}[1]{\mathrm{E}[#1]}
\newcommand{\Var}[1]{\mathrm{Var}[#1]}
\newcommand{\Cov}[2]{\mathrm{Cov}[#1,#2]}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[The vanishing redshift as an EoR probe]{The vanishing redshift as a probe of reionization}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[K. Moriwaki et al.]{
Kana Moriwaki,$^{1}$\thanks{E-mail: kana.moriwaki@utap.phys.s.u-tokyo.ac.jp}
Angus Beane,$^{2}$\thanks{E-mail: angus.beane@cfa.harvard.edu}
Adam Lidz,$^{3}$
and others
\\
% List of institutions
$^{1}$Department of Physics, The University of Tokyo, 7-3-1 Hongo, Bunkyo, Tokyo 113-0033, Japan\\
$^{2}$Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
$^{3}$Department of Physics \& Astronomy, University of Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104, USA\\
% $^{2}$Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, NY 10010, USA\\
}

@@

% Abstract of the paper
\begin{abstract}
The early stages of the Epoch of Reionization (EoR), probed by the 21~cm line,
are sensitive to the detailed formation of the first galaxies. An important
transition occurs on large scales at $z\sim10$, when the 21~cm field changes
from being positively correlated with the density field to being negatively
correlated. The redshift at which this occurs (the ``vanishing redshift'') is
in principle measurable if the 21~cm field can be cross-correlated with a
tracer of the density field (such as galaxies or a line intensity map) at that
high redshift. Since the vanishing redshift is derived from a
cross-correlation, it naively depends on the structure of the 21~cm field and
the tracer field. However, we show that on large scales the transition
redshift is independent of the structure of the tracer field. We verify this
in simulations by assuming a simple model for a mock Ly$\alpha$ intensity
mapping survey. By considering a joint measurement between the Square
Kilometre Array and the Cosmic Dawn Intensity Mapper, we show that the
vanishing redshift is measurable up until about $z\sim x$. Being derived from
a cross-correlation, the vanishing redshift is less impacted by foreground
contamination currently inhibiting 21~cm auto-spectrum measurements.
Therefore, the vanishing redshift is an interesting probe of the early stages
of the EoR, and its detectability and interpretability deserve further study.
\end{abstract}

@@

Because of the difficulties in making 21~cm auto-spectrum measurements, the prospect of measuring 21~cm in cross-correlation with

@@

This cross-correlation was first studied with the CO rotational lines
\citep{2011ApJ...741...70L}, and then the [\CII] \citep{2012ApJ...745...49G}
and Ly$\alpha$ lines \citep{2013ApJ...763..132S}. The possibility of using the
cross-bispectrum the 21~cm and [\CII] fields has been explored
\citep{2018ApJ...867...26B}, along with combining the three cross-spectra with
a 21~cm, [\CII], and [\OIII] overlap \citep{2019ApJ...874..133B}. Because the
vanishing redshift is typically $\zv \sim 10$, it is unclear how much
metal enrichment will have occurred by this early time.

In the case of either galaxies or emission lines, it is expected that the two
are positively correlated at the very beginning of the EoR (though the
emission lines may be very weak due to the lack of significant metal
enhancement). This is because the densest regions host the most mass and
therefore 21~cm emission and galaxy populations. However, the densest regions
will also ionize first {\bf cite something}, and after some time 21~cm and
galaxies will become negatively correlated. We refer to the redshift at which
this transition occurs at a fixed $k$ as the ``vanishing redshift''
$\zv(k)$, assuming that only one such zero exists. In other words, the
cross-correlation at $k$ vanishes at $z=\zv(k)$.

In this work we investigate to what extent the vanishing redshift is
dependent upon the galaxy sample used to define it. We assume a galaxy catalog
selected using an [\OIII] emitter survey. In particular, we are interested in
how \zv{} changes as we vary the parameters of the [\OIII] luminosity-mass
relation for a fixed halo catalog.

\section{Methods} \label{sec:methods}
\subsection{The vanishing redshift} \label{ssec:ztran_def}
The vanishing redshift $\zv(k)$ is defined by a field $x$ which traces the
density field on large scales and the condition that
\beq \label{eq:ztran_def}
P_{21,x}(k, \zv(k)) \equiv 0\text{.}
\eeq
If more than one such \zv{} exists, then we choose the lowest one where
$P_{21,x}$ last transitioned from positive to negative.

@@

\section{Results} \label{sec:results}
\subsection{The vanishing redshift} \label{ssec:ztran_results}

\begin{figure*}
\includegraphics[width=\textwidth]{plots/aveTb_nf.pdf}
\caption{A summary of the ionization and heating history of our three EoR
models: \hot{}, \fid{}, and \cold{}. \textit{Upper}: the ionization history of
each model, showing that reionization begins ($\avg{x_{\HII}}\sim0.01$) at
$z\sim16$ and ends at $z\sim6$ in each model. In detail the ionization
histories are slightly different, but not drastically. \textit{Lower}: the
heating history of each model, showing the strong differences in the monopole
signal. The vanishing redshift \zv{} (defined where $P_{21,\delta}(\zv{}) =
0$) for each model is shown as a vertically dashed line. We see that models in
which heating is more efficient have an earlier \zv{}.}
\label{fig:aveTb_nf}
\end{figure*}

@@

We first investigate the dependence of the vanishing redshift as defined by
the halo field. Fig.~\ref{fig:ztran_vs_k} shows the vanishing redshift
defined by the 21~cm\,--\,density cross-power spectrum (see
Eq.~\ref{eq:ztran_def}) as a function of $k$. The three models considered in
this work are shown in orange, gray, and teal (\hot{}, \fid{}, and \cold{},
respectively).

Fig.~\ref{fig:ztran_vs_k} shows that \zv{} is only weakly a function of $k$.

\begin{figure}
\includegraphics[width=\columnwidth]{plots/ztran_vs_k.pdf}
\caption{The vanishing redshift ($\zv$) as defined by cross-correlating with the
density field (solid lines) and all resolved halos (dashed lines, $M\gtrsim
5\times10^9\,\Msun$), for our three different EoR models (\hot{}, \fid{}, and
\cold{}). We see that \zv{} is only a mild function of $k$ for the three
models, with slightly stronger dependence as the heating of the model
decreases. The most dependence comes in the \cold{} model, where \zv{}
changes by $\sim7\,\%$ from $k=0.1$ to $0.6\,\hoverMpc$. {\bf Some words for
why cold and fid cut off at some k}}
\label{fig:ztran_vs_k}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{plots/ztran_vs_Mcut.pdf}
\caption{The vanishing redshift (\zv) as defined by cross-correlating with
the halo field for our three different EoR models (\hot{}, \fid{}, and
\cold{}) as a function of the cut-off mass. The cross-correlation is only made
between the 21~cm field and all halos with masses $M>M_{\text{cut}}$. One can
see that for the \fid{} and \cold{} models, \zv{} is only mildly a function
of $M_{\text{cut}}$ below a certain cut-off mass ($\sim10^{10.5}\,\Msun$ in
both models). In other words, as long as a galaxy survey is deep enough so as
to resolve a sufficient number of galaxies, the \zv{} is not dependent on the
depth of the survey. The \hot{} model shows more dependence, but this may be
due to sampling noise due to the smaller number of such halos at the higher
\zv{} typical for the \hot{} model.}
\end{figure}

\subsection{Line-intensity mapping} \label{ssec:lim}
We now turn to understanding how \zv{} depends on the parameters of our
generic line-intensity mapping field: the slope ($\alpha$) and lognormal
scatter ($\sigma$). See \S~\ref{ssec:oiii_int_map} for more details of the
construction of the field.

In Fig.~\ref{fig:ztran_vs_alpha_sigma} we plot \zv{} as a function of each of
these parameters\,--\,$\alpha$ and $\sigma$. We see that \zv{} is remarkably
insensitive to the value of either of these parameters, with some small
deviation as a function of $\alpha$. The \hot{}, \fid{}, and \cold{} models
only vary by a total of $1.2\%$, $0.56\%$, $0.52\%$ with respect to $\alpha$,
and $0.070\%$, $0.10\%$, and $0.11\%$ with respect to $\sigma$, respectively.

\begin{figure*}
\includegraphics[width=\textwidth]{plots/ztran_vs_alpha_sigma.pdf}
\caption{The vanishing redshift (\zv) defined by cross-correlating with a
generic two-parameter intensity mapping field. The parameter $\alpha$ sets the
slope of the luminosity-mass relation and the parameter $\sigma$ sets the
lognormal scatter in this relation (see \S~\ref{ssec:oiii_int_map}). The
\hot{}, \fid{}, and \cold{} models only vary by a total of $1.2\%$, $0.56\%$,
$0.52\%$ with respect to $\alpha$, and $0.070\%$, $0.10\%$, and $0.11\%$ with
respect to $\sigma$, respectively. Clearly \zv{} varies only with the
structure of the 21~cm field, and is thus an excellent probe of reionization.}
\label{fig:ztran_vs_alpha_sigma}
\end{figure*}

\section{Detectability}
Up to this point, we have not been specific about which galactic emission line
one might consider using to measure \zv{}\,--\,in fact, our entire formalism
could be executed using a galaxy survey, if one with sufficient redshift
accuracy and depth could be executed. While it will not be possible for a
traditional galaxy survey to extend to $z\sim10$ in the near future, it may be
possible to execute a survey of line-emitting galaxies, e.g., with
[\OIII]-emitters \citep{2018MNRAS.481L..84M, 2019arXiv190610863M}.

@@

The number of modes is given by,
\beq \label{eq:nmodes}
N_m = \frac{4\pi k^2 \delta k}{V_{\text{fund}}}\text{,}
\eeq
where $\delta k$ is the width of the $k$-bin and $V_{\text{fund}}$ is the
volume of a fundamental mode. We assume a square survey area, and so this is
given by,
\beq \label{eq:fund}
V_{\text{fund}} = \frac{(2 \pi)^3}{L_{\perp}^2 L_{\parallel}}\text{,}
\eeq
where $L_{\perp}$ is the  and $L_{\parallel}$ is the comoving length of the
redshift extent of the bin. The optimal redshift binning is not entirely
clear, since the observational target is not the power spectrum. We assume a
redshift bin of $\Delta z = 0.03$, which is smaller than the expected error we
compute on \zv{}, and so is not an unreasonable choice. We assume a $k$-bin
of fairly large width, from $k=0.1$ to $0.4\,\Mpc^{-1}$.

By expanding $P_{21,\Ha}$ as a linear function about \zv{}, we can estimate the
error on \zv{} by
\beq \label{eq:error_on_zst}
\Var{\zv} = \left(\left.\frac{\partial P_{21,\Ha}}{\partial z}\right|_{\zv}\right)^{-2} \Var{P_{21,\Ha}}\text{,}
\eeq
where ${\partial P_{21,\Ha}}/{\partial z}$ is computed numerically.
